Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-05-16 article-models-ttinternal.tex
shortened the long proof

\begin{proof}
The proof is a rather long computation. It was found by starting with the
application of \autoref{eq:extalg-eq1} and making the computation in both ways
from there. The judgmental equality below where \autoref{eq:extalg-eq1} is applied,
is marked with an $*$. The computation goes as follows:
\begin{align*}
\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{Q}}
& \jdeq
  \subst{\epsilon_0}
    {\ctxwk{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{Q}}}
  \\
& \jdeq
  \subst
    {\idtm{\jcomp{}{\epsilon_0}{P}}}
    { \ctxwk
        {\jcomp{}{\epsilon_0}{P}}{\subst{\epsilon_0}
        {\ctxwk{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{Q}}}}}
  \\
& \jdeq
  \subst
    {\idtm{\jcomp{}{\epsilon_0}{P}}}
    { {\ctxwk{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}}
      {\ctxwk{\jcomp{}{\epsilon_0}{P}}{{\ctxext{A}{P}}{\jcomp{}{\epsilon_0}{Q}}}}}
  \\
& \jdeq
  \subst
    {\idtm{\jcomp{}{\epsilon_0}{P}}}
    { {\ctxwk{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}}
      {\ctxwk{\ctxext{{A}{P}}{\jcomp{}{\epsilon_0}{P}}}{\jcomp{}{\epsilon_0}{Q}}}}
  \\
& \jdeq
  \subst
    {\idtm{\jcomp{}{\epsilon_0}{P}}}
    { {\ctxwk{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}}
      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}}
  \\
& \jdeq
  \subst
    {\unfold{\jvcomp{\jcomp{}{\epsilon_0}{P}}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}}
    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}
  \\
& \jdeq
  \subst
    {\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}
    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}
  \\
& \jdeq
  \jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}
  \\
& \jdeq
  \jcomp{}{\jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\epsilon_0}}{Q}
  \\
& \stackrel{*}{\jdeq}
  \jcomp{}{\jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{\epsilon_0}}{Q}
  \\
& \jdeq
  \jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{
    \jcomp{}{\epsilon_0}{Q}}
  \\
& \jdeq
  \subst
    {\tmext{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{\idtm{A}}}{\epsilon_1}}
    {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}
  \\
& \jdeq
  \subst
    {\epsilon_1}
    { {\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{\idtm{A}}}
      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}
      }
  \\
& \jdeq
  \subst
    {\epsilon_1}
    { {\cprojfstf{A}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}}
      {\ctxwk{\ctxext{A}{{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}
      }
  \\
& \jdeq
  \subst
    {\epsilon_1}
    {\jcomp{}{\cprojfstf{A}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}}
  \\
& \jdeq
  \subst
    {\epsilon_1}
    {\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{\jcomp{}{\epsilon_0}{Q}}}
  \\
& \jdeq
  \jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{Q}}.
  \qedhere
\end{align*}
\end{proof}

>>>>

\begin{proof}
We only prove the first inference rule. Let $\jfam{{\Gamma}{A}}{Q}$ be a family.
Then we have the judgmental equalities
\begin{align*}
\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{Q}}
& \jdeq
  \jcomp{}{\idtm{\jcomp{}{\epsilon_0}{P}}}{\jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{Q}}}
  \tag{by \autoref{idfunc-precomp}}
  \\
& \jdeq
  \jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\jcomp{}{\epsilon_0}{Q}}
  \tag{by \autoref{lem:composition-threesome}}
  \\
& \jdeq
  \jcomp{}{\jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\epsilon_0}}{Q}
  \tag{by \autoref{lem:jcomp-jcomp}}
  \\
& \jdeq
  \jcomp{}{\jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{\epsilon_0}}{Q}
  \tag{by \autoref{eq:extalg-eq1}}
  \\
& \jdeq
  \jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{\jcomp{}{\epsilon_0}{Q}}
  \tag{by \autoref{lem:jcomp-jcomp}}
  \\
& \jdeq
  \jcomp{}{\epsilon_1}{\jcomp{}{\idtm{A}}{\jcomp{}{\epsilon_0}{Q}}}
  \tag{by \autoref{lem:composition-threesome}}
  \\
& \jdeq
  \jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{Q}}.
  \tag{by \autoref{idfunc-precomp}}
\end{align*}
\end{proof}
